---
title: "Chapter 8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Applied Exercises
7. In the lab, we applied random forests to the Boston data using mtry=6 and using ntree=25 and ntree=500. Create a plot displaying the test error resulting from random forests on this data set for a more comprehensive range of values for mtry and ntree. You can model your plot after Figure 8.10. Describe the results obtained.
```{r}
library(MASS)
library(randomForest)
```
```{r}
set.seed(1)
train <- sample(1:nrow(Boston), nrow(Boston) / 2)
Boston.train <- Boston[train, -14]
Boston.test <- Boston[-train, -14]
Y.train <- Boston[train, 14]
Y.test <- Boston[-train, 14]
rf.boston1 <- randomForest(Boston.train, y = Y.train, xtest = Boston.test, ytest = Y.test, mtry = ncol(Boston) - 1, ntree = 500)
rf.boston2 <- randomForest(Boston.train, y = Y.train, xtest = Boston.test, ytest = Y.test, mtry = (ncol(Boston) - 1) / 2, ntree = 500)
rf.boston3 <- randomForest(Boston.train, y = Y.train, xtest = Boston.test, ytest = Y.test, mtry = sqrt(ncol(Boston) - 1), ntree = 500)
plot(1:500, rf.boston1$test$mse, col = "green", type = "l", xlab = "Number of Trees", ylab = "Test MSE", ylim = c(10, 19))
lines(1:500, rf.boston2$test$mse, col = "red", type = "l")
lines(1:500, rf.boston3$test$mse, col = "blue", type = "l")
legend("topright", c("m = p", "m = p/2", "m = sqrt(p)"), col = c("green", "red", "blue"), cex = 1, lty = 1)
```
We may see that the Test MSE is very high for a single tree, it decreases as the number of trees increases. Also the Test MSE for all predictors is higher than for half the predictors or the square root of the number of predictors.
## 8. In the lab, a classification tree was applied to the Carseats data set after converting Sales into a qualitative response variable. Now we will seek to predict Sales using regression trees and related approaches, treating the response as a quantitative variable.
a. Split the data set into a training set and a test set.
```{r}
library(ISLR)
set.seed(1)
train = sample(1:nrow(Carseats), nrow(Carseats) / 2)
Car.train = Carseats[train, ]
Car.test = Carseats[-train,]
```
b. Fit a regression tree to the training set. Plot the tree, and interpret the results. What test MSE do you obtain?
```{r}
library(tree)
reg.tree = tree(Sales~.,data = Carseats, subset=train)
reg.tree = tree(Sales~.,data = Car.train)
#Both above formulas outcome the same result.
summary(reg.tree)
```
```{r}
plot(reg.tree)
text(reg.tree ,pretty =0)
```
```{r}
yhat = predict(reg.tree,newdata = Car.test)
mean((yhat - Car.test$Sales)^2)
```
We conclude that the Test MSE is about 4.9.
c. Use cross-validation in order to determine the optimal level of tree complexity. Does pruning the tree improve the test MSE?
```{r}
set.seed(1)
cv.car = cv.tree(reg.tree)
plot(cv.car$size, cv.car$dev, type = "b")
```
In this case, the tree of size 8 is selected by cross-validation. We now prune the tree to obtain the 8-node tree.
```{r}
prune.car = prune.tree(reg.tree, best = 8)
plot(prune.car)
text(prune.car,pretty=0)
```
```{r}
yhat=predict(prune.car, newdata= Car.test)
mean((yhat-Car.test$Sales)^2)
```
We see that pruning the tree increases the Test MSE to 5.1.
d. Use the bagging approach in order to analyze this data. What test MSE do you obtain? Use the importance() function to determine which variables are most important.
```{r}
library(randomForest)
set.seed(1)
bag.car = randomForest(Sales~.,data=Car.train,mtry = 10, importance = TRUE)
```
```{r}
yhat.bag = predict(bag.car,newdata=Car.test)
mean((yhat.bag-Car.test$Sales)^2)
```
```{r}
importance(bag.car)
```
```{r}
varImpPlot(bag.car)
```
The most important variables are the price that company charges for car seats at each site and the quality of the shelving location for the car seats at each site. The test MSE associated with the bagegd regression tree iss 2.55, almost half that obtained using an optimally-pruned single tree.
e. Use random forests to analyze this data. What test MSE do you obtain? Use the importance() function to determine which variables are most important. Describe the effect of m, the number of variables considered at each split, on the error rate obtained.
```{r}
library(randomForest)
set.seed(1)
rf.car = randomForest(Sales~.,data=Car.train,mtry = 3, importance = TRUE)
yhat.rf = predict(rf.car,newdata=Car.test)
mean((yhat.rf-Car.test$Sales)^2)
```
The test set MSE is 2.9; this indicates that random forests doesn’t yield an improvement over bagging in this case.
## 9. This problem involves the OJ data set which is part of the ISLR package.
a. Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.
```{r}
library(ISLR)
set.seed(1)
train = sample(dim(OJ)[1],800)
OJ.train = OJ[train,]
OJ.test = OJ[-train,]
```
b. Fit a tree to the training data, with Purchase as the response and the other variables as predictors. Use the summary() function to produce summary statistics about the tree, and describe the results obtained. What is the training error rate? How many terminal nodes does the tree have?
```{r}
OJ.tree = tree(Purchase~., data=OJ.train)
summary(OJ.tree)
```
The fitted tree has 8 terminal nodes and a training error rate of 0.1588.
c. Type in the name of the tree object in order to get a detailed text output. Pick one of the terminal nodes, and interpret the information displayed.
```{r}
OJ.tree
```
I pick the node labelled 8, which is a terminal node because of the asterisk. The split criterion is LoyalCH < 0.035, the number of observations in that branch is 59 with a deviance of 10.14 and an overall prediction for the branch of MM. Less than x% of the observations in that branch take the value of CH, and the remaining x% take the value of MM.
d. Create a plot of the tree, and interpret the results.
```{r}
plot(OJ.tree)
text(OJ.tree,pretty=TRUE)
```
The most important indicator of “Purchase” appears to be “LoyalCH”, since the first branch differentiates the intensity of customer brand loyalty to CH. In fact, the top three nodes contain “LoyalCH”.
e. Predict the response on the test data, and produce a confusion matrix comparing the test labels to the predicted test labels. What is the test error rate?
```{r}
tree.pred = predict(OJ.tree, newdata = OJ.test, type = "class")
table(tree.pred,OJ.test$Purchase)
```
```{r}
(160+64)/270
```
83% of the test observations are correctly classified so the test error rate is 17%.
f. Apply the cv.tree() function to the training set in order to determine the optimal tree size.
```{r}
cv.OJ = cv.tree(OJ.tree, FUN = prune.misclass)
cv.OJ
```
g. Produce a plot with tree size on the x-axis and cross-validated classification error rate on the y-axis.
```{r}
plot(cv.OJ$size,cv.OJ$dev,type='b', xlab = "Tree size", ylab = "Deviance")
```
h. Which tree size corresponds to the lowest cross-validated classification error rate?
We might see that the 5-node tree is the smallest tree with the lowest classification error rate.
i. Produce a pruned tree corresponding to the optimal tree size obtained using cross-validation. If cross-validation does not lead to selection of a pruned tree, then create a pruned tree with five terminal nodes.
```{r}
prune.OJ = prune.misclass(OJ.tree, best=5)
plot(prune.OJ)
text(prune.OJ,pretty=0)
```
j. Compare the training error rates between the pruned and unpruned trees. Which is higher?
```{r}
tree.pred = predict(prune.OJ, newdata = OJ.test, type = "class")
table(tree.pred,OJ.test$Purchase)
```
```{r}
(160+66)/270
```
In this case, the pruning has the same test error, but it produced a way more interpretable tree.
## 10. We now use boosting to predict Salary in the Hitters data set.
a. Remove the observations for whom the salary information is unknown, and then log-transform the salaries.
```{r}
Hitters = na.omit(Hitters)
Hitters$Salary = log(Hitters$Salary)
```
b. Create a training set consisting of the first 200 observations, and a test set consisting of the remaining observations.
```{r}
train = 1:200
hitters.train = Hitters[train,]
hitters.test = Hitters[-train,]
```
c. Perform boosting on the training set with 1,000 trees for a range of values of the shrinkage parameter λ. Produce a plot with different shrinkage values on the x-axis and the corresponding training set MSE on the y-axis.
```{r}
library(gbm)
```
```{r}
set.seed(1)
pows = seq(-10, -0.2, by = 0.1)
lambdas = 10^pows
train.err = rep(NA, length(lambdas))
for (i in 1:length(lambdas)) {
    boost.hitters = gbm(Salary ~ ., data = hitters.train, distribution = "gaussian", n.trees = 1000, shrinkage = lambdas[i])
    pred.train = predict(boost.hitters, hitters.train, n.trees = 1000)
    train.err[i] = mean((pred.train - hitters.train$Salary)^2)
}
plot(lambdas, train.err, type = "b", xlab = "Shrinkage values", ylab = "Training MSE")
```
d. Produce a plot with different shrinkage values on the x-axis and the corresponding test set MSE on the y-axis.
```{r}
set.seed(1)
test.err <- rep(NA, length(lambdas))
for (i in 1:length(lambdas)) {
    boost.hitters = gbm(Salary ~ ., data = hitters.train, distribution = "gaussian", n.trees = 1000, shrinkage = lambdas[i])
    yhat = predict(boost.hitters, hitters.test, n.trees = 1000)
    test.err[i] = mean((yhat - hitters.test$Salary)^2)
}
plot(lambdas, test.err, type = "b", xlab = "Shrinkage values", ylab = "Test MSE")
```
```{r}
min(test.err)
```
```{r}
lambdas[which.min(test.err)]
```
e.Compare the test MSE of boosting to the test MSE that results from applying two of the regression approaches seen in Chapters 3 and 6.
```{r}
library(glmnet)
```
```{r}
fit1 = lm(Salary ~ ., data = hitters.train)
pred1 = predict(fit1, hitters.test)
mean((pred1 - hitters.test$Salary)^2)
```
```{r}
x = model.matrix(Salary ~ ., data = hitters.train)
x.test = model.matrix(Salary ~ ., data = hitters.test)
y = hitters.train$Salary
fit2 = glmnet(x, y, alpha = 0)
pred2 = predict(fit2, s = 0.01, newx = x.test)
mean((pred2 - hitters.test$Salary)^2)
```
The test MSE for boosting is lower than for linear regression and ridge regression.
f. Which variables appear to be the most important predictors in the boosted model?
```{r}
boost.hitters <- gbm(Salary ~ ., data = hitters.train, distribution = "gaussian", n.trees = 1000, shrinkage = lambdas[which.min(test.err)])
summary(boost.hitters)
```
“CAtBat” is by far the most important variable.
g. Now apply bagging to the training set. What is the test set MSE for this approach?
```{r}
set.seed(1)
bag.hitters <- randomForest(Salary ~ ., data = hitters.train, mtry = 19, ntree = 500)
yhat.bag <- predict(bag.hitters, newdata = hitters.test)
mean((yhat.bag - hitters.test$Salary)^2)
```
The test MSE for bagging is 0.23, which is slightly lower than the test MSE for boosting.
## 11. This question uses the Caravan data set.
a. Create a training set consisting of the first 1,000 observations, and a test set consisting of the remaining observations.
```{r}
train = 1:1000
Caravan$Purchase = ifelse(Caravan$Purchase == "Yes", 1, 0)
Caravan.train = Caravan[train,]
Caravan.test = Caravan[-train,]
```
b. Fit a boosting model to the training set with Purchase as the response and the other variables as predictors. Use 1,000 trees, and a shrinkage value of 0.01. Which predictors appear to be the most important?
```{r}
set.seed(1)
boost.caravan = gbm(Purchase ~ ., data = Caravan.train, distribution = "gaussian", n.trees = 1000, shrinkage = 0.01)
```
```{r}
summary(boost.caravan)
```
The variables “PPERSAUT” and “MKOOPKLA” are the two most important variables.
c. Use the boosting model to predict the response on the test data. Predict that a person will make a purchase if the estimated prob- ability of purchase is greater than 20 %. Form a confusion ma- trix. What fraction of the people predicted to make a purchase do in fact make one? How does this compare with the results obtained from applying KNN or logistic regression to this data set?
```{r}
probs.test <- predict(boost.caravan, Caravan.test, n.trees = 1000, type = "response")
pred.test <- ifelse(probs.test > 0.2, 1, 0)
table(Caravan.test$Purchase, pred.test)
```
For boosting, the fraction of people predicted to make a purchase that in fact make one is 0.2156863.
```{r}
logit.caravan <- glm(Purchase ~ ., data = Caravan.train, family = "binomial")
```
```{r}
probs.test2 <- predict(logit.caravan, Caravan.test, type = "response")
```
```{r}
pred.test2 <- ifelse(probs.test > 0.2, 1, 0)
table(Caravan.test$Purchase, pred.test2)
```
For logistic regression, the fraction of people predicted to make a purchase that in fact make one is again 0.2156863.
